---
title: "Preprocess phenotype data"
output: html_document
date: "2024-01-03"
---

```{r global_chunk_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, warning = FALSE, message = FALSE, cache = TRUE, cache.lazy = FALSE, fig.width = 8, fig.height = 4.5)

here::i_am("submission/draft/preprocess_phenodata.Rmd")
```

```{r, echo = TRUE}
library(tidyverse)
library(readxl)
library(here)
library(janitor)
library(fs)
```

# Functions

```{r}
# Divides Excel sector map into sub-grids, and then converts those into long format
convert_sectormap_to_long <- function(sectormap) {
  sectormat <- as.matrix(sectormap)
  
  nr <- nrow(sectormat)
  nc <- ncol(sectormat)
  
  heading_locs <- str_detect(sectormat, "^Sector") %>% which
  heading_rows <- heading_locs %% nr
  heading_cols <- floor(heading_locs / nr) + 1
  
  core_map <- lapply(1:length(heading_rows), function(i) {
    start_row <- heading_rows[i]
    start_col <- heading_cols[i]
    
    end_row <- min(heading_rows[heading_rows > start_row] - 1, nr, na.rm = TRUE)
    end_col <- min(heading_cols[heading_cols > start_col] - 1, nc, na.rm = TRUE)
    
    sector_label <- str_extract(sectormat[start_row, start_col], "[0-9]+")
    
    df <- sectormat[start_row:end_row,start_col:end_col] %>% 
      as_tibble %>% 
      row_to_names(row_number = 1) 
    colnames(df)[1] <- "row"
    
    all_na_cols <- sapply(df, function(x) all(is.na(x)))
    
    df <- df[,!all_na_cols]
    
    df %>%
      gather(col, core, -row) %>%
      filter(!is.na(core)) %>%
      mutate(core = as.numeric(core), sector = as.numeric(sector_label)) %>%
      arrange(row, col)
  }) %>% bind_rows()
  
  return(core_map)
}
```

# Clinical data

```{r}
clinical_data_file <- here::here("data", "raw", "bc-1508_tma_clin_data_2023-08-07.xlsx")
```

```{r}
clinical_data <- read_excel(clinical_data_file, sheet = "bc-1508", na = c("", "NA", "n/a", "Unk"))

# Clinical data table contains some invisible characters
clinical_data <- clinical_data %>%
  mutate(acc_num=replace_non_ascii(acc_num),
         study_id=replace_non_ascii(study_id)) %>%
  ## Remove leading zeros after hyphens, these are inconsistently formatted in both tables :(
  mutate(acc_num=str_replace(acc_num, "-(0+)", "-")) %>%
  mutate(age_dx=as.numeric(age_dx),
         age_surg=as.numeric(age_surg),
         tx=str_replace(tx, " only", "")) %>%
  mutate(
    chemo = grepl("chemo", tx),
    rt = grepl("RT", tx),
    brachy = grepl("brachy", tx)
  ) %>%
  mutate(
    ## Temporary solution, of no relevance as p53abn is the focus anyways
    eclass2_ngs=dplyr::recode(eclass2_ngs, "discrepant: MMRd, NSMP/p53wt"="NSMP/p53wt"),
    eclass2_pt=dplyr::recode(eclass2_pt, "discrepant: MMRd, NSMP/p53wt"="NSMP/p53wt"),
    stage_main=str_extract(stage_full, "I*V*")
  )
```

```{r}
clinical_output_path <- here::here("data", "processed", "clinical_data.tsv")

write_tsv(clinical_data, file = clinical_output_path)
```


# B & T data

## Identifier wrangling

Resolving inconsistencies between ID schemes ...

```{r}
bt_dir <- here::here("data", "raw", "B&T Panel")
bt_22004_mapfile <- file.path(bt_dir, "22-004_revised_map.xlsx")
bt_datafile <- file.path(bt_dir, "FINAL all TMAs B+T cell densities 22JUN2023- do not edit.xlsx")

bt_22004_map <- read_excel(bt_22004_mapfile, sheet = "Core ID", skip = 2)

sectormaps <- list(
  '22-004A'=read_excel(bt_22004_mapfile, sheet = "SectorMap A - BnT", col_names = FALSE, skip = 2),
  '22-004B'=read_excel(bt_22004_mapfile, sheet = "SectorMap B - BnT", col_names = FALSE, skip = 2)
)

core_ids <- list(
  '22-004A'=read_excel(bt_22004_mapfile, sheet = "Core ID", skip = 2),
  '22-004B'=read_excel(bt_22004_mapfile, sheet = "Core ID", skip = 2)
)
```

```{r}
sectormaps_long <- lapply(names(sectormaps), function(x) {
  convert_sectormap_to_long(sectormaps[[x]]) %>%
    mutate(tma=x) %>%
    select(tma, everything())
}) %>% bind_rows()

core_ids_long <- lapply(names(core_ids), function(x) {
  df <- core_ids[[x]] %>%
    mutate(tma=x) %>%
    rename(core=CoreID) %>%
    select(tma, everything())
  colnames(df) <- str_replace(colnames(df), "( )*#$", "")
  return(df)
}) %>% bind_rows()

# Sample map, in long format, derived from the weird TMA maps
sample_map_22004 <- sectormaps_long %>% 
  inner_join(core_ids_long) %>%
  rename(acc_num=Accession, block=`Block ID(s)`, backup=`Spare Block(s)`) %>%
  select(tma, sector, core, row, col, acc_num, block, backup) %>%
  dplyr::mutate(
    row = as.character(row),
    col = as.character(col)
  )
```

```{r}
bt_sheetnames <- excel_sheets(bt_datafile)

bt_data <- lapply(bt_sheetnames, function(sn) {
  df <- read_excel(bt_datafile, na = c("", "x", "NA", "n/a", "x", "n/a- Pathologist annotated"), sheet = sn)
  df <- df %>%
    dplyr::mutate(tma=str_extract(`Image Tag`, "^[0-9]{2}\\-[0-9]{3}( V1B1 [AB])?")) %>%
    dplyr::mutate(tma=str_replace(tma, " V1B1 ", "")) %>%
    ## The labeling scheme seems to be different each time ...
    dplyr::mutate(
      secrowcol=str_extract(`Image Tag`, "(?<=Core\\[)[0-9A-Z,]+(?=\\])")
    ) %>%
    separate(secrowcol, sep = ",", c("sector", "row2", "col2")) %>%
    dplyr::mutate(
      sector = as.numeric(sector),
      # Rows and columns are reversed
      row = col2,
      col = row2
    ) %>%
    dplyr::select(-c(row2, col2))
  
  return(df)
}) 
names(bt_data) <- bt_sheetnames

bt_data_merged <- reduce(bt_data, dplyr::inner_join) %>%
  dplyr::rename(acc_num=`Accession#`,
                block=`Block#`)

bt_data_merged_no22004 <- bt_data_merged %>%
  filter(!str_detect(tma, "^22\\-004"))
bt_data_merged_22004 <- bt_data_merged %>%
  filter(str_detect(tma, "^22\\-004")) %>%
  select(-c(acc_num, block)) %>%
  left_join(sample_map_22004) %>%
  select(-c(core, backup))


bt_data_all <- bt_data_merged_no22004 %>%
  bind_rows(bt_data_merged_22004) %>%
  dplyr::select(tma, sector, row, col, acc_num, everything()) %>%
  dplyr::arrange(tma, sector, row, col, acc_num) %>%
  dplyr::mutate(
    acc_num=str_replace(acc_num, "-(0+)", "-")
  ) %>%
  # THIS WILL NEED TO GO IN B&T ANALYSIS TOO
  dplyr::mutate(
    acc_num=dplyr::recode(
      acc_num,
      'O3S-2724'='VS03-2724',
      'YW-EC002'='VS08-19857',
      'YW-EC003'='VS10-497',
      'YW-EC005'='VS10-25417',
      'YW-EC008'='VS13-5723',
      'YW-EC010'='VS13-19793',
      'YW-EC014'='VS16-22196'
    )
  )
```

## Filtering

```{r}
colname_map <- c(
  'Notes'='notes',
  '% tumour area2'='area_tumour_pct',
  '% stroma area2'='area_stroma_pct',
  "tissue area (mm²)"='area_tissue_mm',
  "tumour area (mm²)"='area_tumour_mm',
  "stroma area (mm²)"='area_stroma_mm',
  "TUMOUR CD20+CD79a+ (cells/mm2)"='dens_bcell_cd20pos_tumour',
  "STROMA CD20+CD79a+ (cells/mm2)"='dens_bcell_cd20pos_stroma',
  "TUMOUR CD20-CD70a+ (cells/mm2)"='dens_bcell_cd20neg_tumour',
  "STROMA CD20-CD70a+ (cells/mm2)"='dens_bcell_cd20neg_stroma',
  "TUMOUR CD3+CD8+FOXP3+ (cells/mm2)"='dens_cd8tcell_foxp3pos_tumour',
  "STROMA CD3+CD8+FOXP3+ (cells/mm2)"='dens_cd8tcell_foxp3pos_stroma',
  "TUMOUR CD3+CD8-FOXP3+ (cells/mm2)"='dens_treg_tumour',
  "STROMA CD3+CD8-FOXP3+ (cells/mm2)"='dens_treg_stroma',
  "TUMOUR CD3+CD8-FOXP3- (cells/mm2)"='dens_thelper_tumour',
  "STROMA CD3+CD8-FOXP3- (cells/mm2)"='dens_thelper_stroma',
  "TUMOUR CD3+CD8+FOXP3- (cells/mm2)"='dens_tcyto_tumour',
  "STROMA CD3+CD8+FOXP3- (cells/mm2)"='dens_tcyto_stroma',
  "TUMOUR CD20+CD79a+2"='count_bcell_cd20pos_tumour',
  "STROMA CD20+CD79a+3"='count_bcell_cd20pos_stroma',
  "TUMOUR CD20-CD70a+4"='count_bcell_cd20neg_tumour',
  "STROMA CD20-CD70a+5"='count_bcell_cd20neg_stroma',
  "TUMOUR CD3+CD8+FOXP3+"='count_cd8tcell_foxp3pos_tumour',
  "STROMA CD3+CD8+FOXP3+"='count_cd8tcell_foxp3pos_stroma',
  "TUMOUR CD3+CD8-FOXP3+"='count_treg_tumour',
  "STROMA CD3+CD8-FOXP3+"='count_treg_stroma',
  "TUMOUR CD3+CD8-FOXP3-"='count_thelper_tumour',
  "STROMA CD3+CD8-FOXP3-2"='count_thelper_stroma',
  "TUMOUR CD3+CD8+FOXP3-12"='count_tcyto_tumour',
  "STROMA CD3+CD8+FOXP3-13"='count_tcyto_stroma'
)

bt_data_labeled <- bt_data_all %>% 
  rename_with(~ colname_map[.], .cols = names(colname_map)) %>%
  filter(ifelse(is.na(notes), TRUE, !str_detect(notes, pattern = "(Exclude|exclude)")),
         !str_detect(acc_num, "LANDMARK")) %>%
  mutate(acc_num=str_replace(acc_num, " \\(.*", "")) %>%
  select(c(tma, sector, row, col, acc_num, block, unname(colname_map)))
```

```{r}
bt_output_path <- here::here("data", "processed", "bt_labeled.tsv")

write_tsv(bt_data_labeled, file = bt_output_path)
```

# Adaptive response panel

## Identifier wrangling 

Resolving inconsistencies between ID schemes ...

```{r}
ar_dir <- here::here("data", "raw", "AdaptR Panel")
ar_map_files <- file.path(ar_dir, "AdaptR Final (minus CD8-PD1)", 
                          c("10-005 G3 Endometrioid Endometrial Carcinoma.xls",
                            "10-006 High-Grade Serous Endometrial Carcinoma.xls",
                            "22-004 P53 Abnormal Endometrial Carcinoma Tumours (mixed sites) - BnT_AR Map.xlsx")
)

sectormaps <- list(
  '10-005'=read_excel(ar_map_files[1], sheet = "SectorMap-AR_SDM", col_names = FALSE, skip = 2),
  '10-006'=read_excel(ar_map_files[2], sheet = "SectorMap-AR-SDM", col_names = FALSE, skip = 2),
  '22-004A'=read_excel(ar_map_files[3], sheet = "SectorMap A-AR-SDM", col_names = FALSE, skip = 2),
  '22-004B'=read_excel(ar_map_files[3], sheet = "SectorMap B-AR-SDM", col_names = FALSE, skip = 2)
)

core_ids <- list(
  '10-005'=read_excel(ar_map_files[1], sheet = "CoreID", skip = 2),
  '10-006'=read_excel(ar_map_files[2], sheet = "CoreID", skip = 2),
  '22-004A'=read_excel(ar_map_files[3], sheet = "Core ID", skip = 2),
  '22-004B'=read_excel(ar_map_files[3], sheet = "Core ID", skip = 2)
)
```



```{r}
sectormaps_long <- lapply(names(sectormaps), function(x) {
  convert_sectormap_to_long(sectormaps[[x]]) %>%
    mutate(tma=x) %>%
    select(tma, everything())
}) %>% bind_rows()

core_ids_long <- lapply(names(core_ids), function(x) {
  df <- core_ids[[x]] %>%
    mutate(tma=x) %>%
    rename(core=CoreID) %>%
    select(tma, everything())
  colnames(df) <- str_replace(colnames(df), "( )*#$", "")
  return(df)
}) %>% bind_rows()

# Sample map, in long format, derived from the weird TMA maps
sample_map <- sectormaps_long %>% 
  inner_join(core_ids_long) %>%
  rename(acc_num=Accession, block=Block, backup=Backup) %>%
  select(tma, sector, core, row, col, acc_num, block, backup) %>%
  dplyr::mutate(
    row = trimws(as.character(row)),
    col = trimws(as.character(col))
  )
```

### 10-005 and 10-006

Note that 10-005 & 6 is labeled differently from 22-004, for some reason. 

```{r}
ar_pheno_100056_file_path <- file.path(ar_dir, "AdaptR Final (minus CD8-PD1)", "FINAL 10-005&6 AR PHENO 23Nov2022.xlsx")
ar_pheno_100056_sheetnames <- excel_sheets(ar_pheno_100056_file_path)

ar_pheno_100056_data <- lapply(ar_pheno_100056_sheetnames, function(sn) {
  df <- read_excel(ar_pheno_100056_file_path, na = c("", "x", "excluded", "n/a"), sheet = sn)
  df <- df %>%
    dplyr::mutate(tma=str_extract(`Image Tag`, "^[0-9]{2}_[0-9]{3}[A-Z]?")) %>%
    ## The labeling scheme seems to be different each time ... 
    dplyr::mutate(
      tma = dplyr::recode(tma,
                          "10_005V"="10-005",
                          "10_006V"="10-006")
    ) %>%
    dplyr::mutate(
      secrowcol=str_extract(`Image Tag`, "(?<=Core\\[)[0-9,]+(?=\\])")
    ) %>%
    separate(secrowcol, sep = ",", c("sector", "row", "col")) %>%
    dplyr::mutate(
      sector = as.numeric(sector)
    )
  
  return(df)
}) 
names(ar_pheno_100056_data) <- ar_pheno_100056_sheetnames

ar_pheno_100056_merged <- reduce(ar_pheno_100056_data, dplyr::inner_join)

ar_pheno_100056_labeled <- ar_pheno_100056_merged %>%
  left_join(sample_map) %>%
  dplyr::select(tma, acc_num, sector, row, col, core, Notes, everything()) %>%
  dplyr::rename(notes=Notes) %>%
  dplyr::arrange(tma, sector, row, col, acc_num) %>%
  dplyr::mutate(
    acc_num=str_replace(acc_num, "-(0+)", "-")
  ) %>%
  dplyr::mutate(
    acc_num=dplyr::recode(
      acc_num,
      'O3S-2724'='VS03-2724',
      'YW-EC002'='VS08-19857',
      'YW-EC003'='VS10-497',
      'YW-EC005'='VS10-25417',
      'YW-EC008'='VS13-5723',
      'YW-EC010'='VS13-19793',
      'YW-EC014'='VS16-22196'
    )
  )
```

```{r}
ar_area_100056_file_path <- file.path(ar_dir, "AdaptR Final (minus CD8-PD1)", "FINAL 10-005&6 ADAPTR AREA QUANT 30JUN23.xlsx")
ar_area_100056_sheetnames <- excel_sheets(ar_area_100056_file_path)


ar_area_100056_data <- lapply(ar_area_100056_sheetnames, function(sn) {
  df <- read_excel(ar_area_100056_file_path, na = c("", "x", "excluded", "n/a"), sheet = sn)
  df <- df %>%
    dplyr::mutate(tma=str_extract(`Image Tag`, "^[0-9]{2}_[0-9]{3}[A-Z]?")) %>%
    ## The labeling scheme seems to be different each time ... 
    dplyr::mutate(
      tma = dplyr::recode(tma,
                          "10_005V"="10-005",
                          "10_006V"="10-006")
    ) %>%
    dplyr::mutate(
      secrowcol=str_extract(`Image Tag`, "(?<=Core\\[)[0-9,]+(?=\\])")
    ) %>%
    separate(secrowcol, sep = ",", c("sector", "row", "col")) %>%
    dplyr::mutate(
      sector = as.numeric(sector)
    )
  
  return(df)
}) 
names(ar_area_100056_data) <- ar_area_100056_sheetnames

ar_area_100056_merged <- reduce(ar_area_100056_data, dplyr::inner_join)

ar_area_100056_labeled <- ar_area_100056_merged %>%
  left_join(sample_map) %>%
  dplyr::select(tma, acc_num, sector, row, col, core, Notes, everything()) %>%
  dplyr::rename(notes=Notes) %>%
  dplyr::arrange(tma, sector, row, col, acc_num) %>%
  dplyr::mutate(
    acc_num=str_replace(acc_num, "-(0+)", "-")
  ) %>%
  dplyr::mutate(
    acc_num=dplyr::recode(
      acc_num,
      'O3S-2724'='VS03-2724',
      'YW-EC002'='VS08-19857',
      'YW-EC003'='VS10-497',
      'YW-EC005'='VS10-25417',
      'YW-EC008'='VS13-5723',
      'YW-EC010'='VS13-19793',
      'YW-EC014'='VS16-22196'
    )
  )
```


#### Filtering

```{r}
colname_map <- c(
  "Tumour Area (mm²)"='area_tumour_mm',
  "Stroma Area (mm²)"='area_stroma_mm',
  "Tumour: CD68+PDL1-IDO1- Cells"='count_mac_pdl1neg_ido1neg_tumour',
  "Stroma: CD68+PDL1-IDO1- Cells"='count_mac_pdl1neg_ido1neg_stroma',
  "Tumour: CD68+PDL1+IDO1- Cells"='count_mac_pdl1pos_ido1neg_tumour',
  "Stroma: CD68+PDL1+IDO1- Cells"='count_mac_pdl1pos_ido1neg_stroma',
  "Tumour: CD68+PDL1+IDO1+ Cells"='count_mac_pdl1pos_ido1pos_tumour',
  "Stroma: CD68+PDL1+IDO1+ Cells"='count_mac_pdl1pos_ido1pos_stroma',
  "Tumour: CD68+PDL1-IDO1+ Cells"='count_mac_pdl1neg_ido1pos_tumour',
  "Stroma: CD68+PDL1-IDO1+ Cells"='count_mac_pdl1neg_ido1pos_stroma',
  "TumourAverage CD8+PD1-"='count_tcyto_pd1neg_tumour',
  "Stroma Average CD8+PD1-"='count_tcyto_pd1neg_stroma',
  "TumourAverage CD8+PD1+"='count_tcyto_pd1pos_tumour',
  "Stroma Average CD8+PD1+"='count_tcyto_pd1pos_stroma',
  "Tumour: CD8-PD1+ Cells"='count_tcd8neg_pd1pos_tumour',
  "Stroma: CD8-PD1+ Cells"='count_tcd8neg_pd1pos_stroma'
)

ar_pheno_100056_labeled <- ar_pheno_100056_labeled %>% 
  rename_with(~ colname_map[.], .cols = names(colname_map)) %>%
  filter(ifelse(is.na(notes), TRUE, !str_detect(notes, pattern = "(Exclude|exclude|discard)"))) %>%
  mutate(acc_num=str_replace(acc_num, " \\(.*", "")) %>%
  select(c(tma, sector, row, col, acc_num, block, unname(colname_map)))
```


```{r}
colname_map_area <- c(
  "Tumour Area (mm²)"='area_tumour_mm',
  "Stroma Area (mm²)"='area_stroma_mm',
  "Tumour: PDL1+IDO1+CD68- Area (μm²)"='area_pdl1pos_ido1pos_cd68neg_tumour',
  "Tumour: PDL1+IDO1-CD68- Area (μm²)"='area_pdl1pos_ido1neg_cd68neg_tumour',
  "Tumour: PDL1-IDO1+CD68- Area (μm²)"='area_pdl1neg_ido1pos_cd68neg_tumour',
  "Tumour: PDL1+IDO1-CD68+ Area (μm²)"='area_pdl1pos_ido1neg_cd68pos_tumour',
  "Tumour: PDL1-IDO1+CD68+ Area (μm²)"='area_pdl1neg_ido1pos_cd68pos_tumour',
  "Tumour: PDL1+IDO1+CD68+ Area (μm²)"='area_pdl1pos_ido1pos_cd68pos_tumour',
  "Tumour: CD68+IDO1-PDL1- Area (μm²)"='area_pdl1neg_ido1neg_cd68pos_tumour',
  "Tumour: PDL1-IDO1-CD68- Area (μm²)"='area_pdl1neg_ido1neg_cd68neg_tumour',
  "Stroma: PDL1+IDO1+CD68- Area (μm²)"='area_pdl1pos_ido1pos_cd68neg_stroma',
  "Stroma: PDL1+IDO1-CD68- Area (μm²)"='area_pdl1pos_ido1neg_cd68neg_stroma',
  "Stroma: PDL1-IDO1+CD68- Area (μm²)"='area_pdl1neg_ido1pos_cd68neg_stroma',
  "Stroma: PDL1+IDO1-CD68+ Area (μm²)"='area_pdl1pos_ido1neg_cd68pos_stroma',
  "Stroma: PDL1-IDO1+CD68+ Area (μm²)"='area_pdl1neg_ido1pos_cd68pos_stroma',
  "Stroma: PDL1+IDO1+CD68+ Area (μm²)"='area_pdl1pos_ido1pos_cd68pos_stroma',
  "Stroma: CD68+IDO1-PDL1- Area (μm²)"='area_pdl1neg_ido1neg_cd68pos_stroma',
  "Stroma: PDL1-IDO1-CD68- Area (μm²)"='area_pdl1neg_ido1neg_cd68neg_stroma'
)

ar_area_100056_labeled <- ar_area_100056_labeled %>% 
  rename_with(~ colname_map_area[.], .cols = names(colname_map_area)) %>%
  filter(ifelse(is.na(notes), TRUE, !str_detect(notes, pattern = "(Exclude|exclude|discard)"))) %>%
  mutate(acc_num=str_replace(acc_num, " \\(.*", "")) %>%
  select(c(tma, sector, row, col, acc_num, block, unname(colname_map_area)))
```


### 22-004

```{r}
ar_pheno_22004_file_path <- file.path(ar_dir, "22-004 A B AdaptR FINAL count data 9NOV2023.xlsx")
ar_pheno_22004_sheetnames <- excel_sheets(ar_pheno_22004_file_path)

ar_pheno_22004_data <- lapply(ar_pheno_22004_sheetnames, function(sn) {
  df <- read_excel(ar_pheno_22004_file_path, na = c("", "x"), sheet = sn)
  df <- df %>%
    dplyr::mutate(tma=str_extract(`Image Tag`, "^[0-9]{2}_[0-9]{3}[A-Z]?")) %>%
    ## The labeling scheme seems to be different each time ... 
    dplyr::mutate(
      tma = dplyr::recode(tma,
                          "22_004A"="22-004A",
                          "22_004V"="22-004B")
    ) %>%
    dplyr::mutate(
      secrowcol=str_extract(`Image Tag`, "(?<=Core\\[)[0-9,]+(?=\\])")
    ) %>%
    separate(secrowcol, sep = ",", c("sector", "row2", "col2")) %>%
    dplyr::mutate(
      sector = as.numeric(sector),
      row = ifelse(tma == "22-004B", col2, row2),
      col = ifelse(tma == "22-004B", row2, col2)
    ) %>%
    dplyr::select(-c(row2, col2))
  
  return(df)
}) 
names(ar_pheno_22004_data) <- ar_pheno_22004_sheetnames

# Manually checked that this has the same number of rows as all the sub-data frames
ar_pheno_22004_merged <- reduce(ar_pheno_22004_data, dplyr::inner_join, by = c("Image Tag", "tma", "sector", "row", "col", "Classified Area (mm²)", "Stroma Area (mm²)", "Tumour Area (mm²)", "% Stroma", "% Tumour"))

ar_pheno_22004_labeled <- ar_pheno_22004_merged %>%
  left_join(sample_map) %>%
  dplyr::select(tma, acc_num, sector, row, col, core, notes.y, everything()) %>%
  dplyr::select(-c(notes.x)) %>%
  dplyr::rename(notes=notes.y) %>%
  dplyr::arrange(tma, sector, row, col, acc_num) %>%
  dplyr::mutate(
    acc_num=str_replace(acc_num, "-(0+)", "-")
  ) %>%
  # THIS WILL NEED TO GO IN B&T ANALYSIS TOO
  dplyr::mutate(
    acc_num=dplyr::recode(
      acc_num,
      'O3S-2724'='VS03-2724',
      'YW-EC002'='VS08-19857',
      'YW-EC003'='VS10-497',
      'YW-EC005'='VS10-25417',
      'YW-EC008'='VS13-5723',
      'YW-EC010'='VS13-19793',
      'YW-EC014'='VS16-22196'
    )
  )
```


```{r}
ar_area_22004_file_path <- file.path(ar_dir, "AdaptR Final (minus CD8-PD1)", "22-004 A B AdaptR FINAL data 30JUNE2023.xlsx")
ar_area_22004_sheetnames <- excel_sheets(ar_area_22004_file_path)

ar_area_22004_data <- lapply(ar_area_22004_sheetnames, function(sn) {
  df <- read_excel(ar_area_22004_file_path, na = c("", "x"), sheet = sn)
  df <- df %>%
    dplyr::mutate(tma=str_extract(`Image Tag`, "^[0-9]{2}_[0-9]{3}[A-Z]?")) %>%
    ## The labeling scheme seems to be different each time ... 
    dplyr::mutate(
      tma = dplyr::recode(tma,
                          "22_004A"="22-004A",
                          "22_004V"="22-004B")
    ) %>%
    dplyr::mutate(
      secrowcol=str_extract(`Image Tag`, "(?<=Core\\[)[0-9,]+(?=\\])")
    ) %>%
    separate(secrowcol, sep = ",", c("sector", "row2", "col2")) %>%
    dplyr::mutate(
      sector = as.numeric(sector),
      row = ifelse(tma == "22-004B", col2, row2),
      col = ifelse(tma == "22-004B", row2, col2)
    ) %>%
    dplyr::select(-c(row2, col2))
  
  return(df)
}) 
names(ar_area_22004_data) <- ar_area_22004_sheetnames

# Manually checked that this has the same number of rows as all the sub-data frames
ar_area_22004_merged <- reduce(ar_area_22004_data, dplyr::inner_join, by = c("Image Tag", "tma", "sector", "row", "col", "Classified Area (mm²)", "Stroma Area (mm²)", "Tumour Area (mm²)", "% Stroma", "% Tumour"))

ar_area_22004_labeled <- ar_area_22004_merged %>%
  left_join(sample_map) %>%
  dplyr::select(tma, acc_num, sector, row, col, core, Notes, everything()) %>%
  dplyr::rename(notes=Notes) %>%
  dplyr::arrange(tma, sector, row, col, acc_num) %>%
  dplyr::mutate(
    acc_num=str_replace(acc_num, "-(0+)", "-")
  ) %>%
  # THIS WILL NEED TO GO IN B&T ANALYSIS TOO
  dplyr::mutate(
    acc_num=dplyr::recode(
      acc_num,
      'O3S-2724'='VS03-2724',
      'YW-EC002'='VS08-19857',
      'YW-EC003'='VS10-497',
      'YW-EC005'='VS10-25417',
      'YW-EC008'='VS13-5723',
      'YW-EC010'='VS13-19793',
      'YW-EC014'='VS16-22196'
    )
  )
```

#### Filtering

```{r}
colname_map <- c(
  "Tumour Area (mm²)"='area_tumour_mm',
  "Stroma Area (mm²)"='area_stroma_mm',
  "Tumour: CD68+PDL1-IDO1- Cells"='count_mac_pdl1neg_ido1neg_tumour',
  "Stroma:CD68+PDL1-IDO1- Cells"='count_mac_pdl1neg_ido1neg_stroma',
  "Tumour: CD68+PDL1+IDO1- Cells"='count_mac_pdl1pos_ido1neg_tumour',
  "Stroma:  CD68+PDL1+IDO1- Cells"='count_mac_pdl1pos_ido1neg_stroma',
  "Tumour: CD68+PDL1+IDO1+ Cells"='count_mac_pdl1pos_ido1pos_tumour',
  "Stroma: CD68+PDL1+IDO1+ Cells"='count_mac_pdl1pos_ido1pos_stroma',
  "Tumour: CD68+PDL1-IDO1+ Cells"='count_mac_pdl1neg_ido1pos_tumour',
  "Stroma: CD68+PDL1-IDO1+ Cells"='count_mac_pdl1neg_ido1pos_stroma',
  "Tumour: CD8+PD1- Cells"='count_tcyto_pd1neg_tumour',
  "Stroma: CD8+PD1- Cells"='count_tcyto_pd1neg_stroma',
  "Tumour: CD8+PD1+ Cells"='count_tcyto_pd1pos_tumour',
  "Stroma: CD8+PD1+ Cells"='count_tcyto_pd1pos_stroma',
  "Tumour: CD8-PD1+ Cells"='count_tcd8neg_pd1pos_tumour',
  "Stroma: CD8-PD1+ Cells"='count_tcd8neg_pd1pos_stroma'
)

ar_pheno_22004_labeled <- ar_pheno_22004_labeled %>% 
  rename_with(~ colname_map[.], .cols = names(colname_map)) %>%
  # This rule shouldn't apply
  filter(ifelse(is.na(notes), TRUE, !str_detect(notes, pattern = "(Exclude|exclude)"))) %>%
  # This rule shouldn't apply
  mutate(acc_num=str_replace(acc_num, " \\(.*", "")) %>%
  select(c(tma, sector, row, col, acc_num, block, unname(colname_map)))
```

```{r}
colname_map_area2 <- c(
  "Tumour Area (mm²)"='area_tumour_mm',
  "Stroma Area (mm²)"='area_stroma_mm',
  "Tumour: PDL1+IDO1+CD68- Area (μm²)"='area_pdl1pos_ido1pos_cd68neg_tumour',
  "Tumour: PDL1+IDO1-CD68- Area (μm²)"='area_pdl1pos_ido1neg_cd68neg_tumour',
  "Tumour: PDL1-IDO1+CD68- Area (μm²)"='area_pdl1neg_ido1pos_cd68neg_tumour',
  "Tumour: PDL1+IDO1-CD68+ Area (μm²)"='area_pdl1pos_ido1neg_cd68pos_tumour',
  "Tumour: PDL1-IDO1+CD68+ Area (μm²)"='area_pdl1neg_ido1pos_cd68pos_tumour',
  "Tumour: PDL1+IDO1+CD68+ Area (μm²)"='area_pdl1pos_ido1pos_cd68pos_tumour',
  "Tumour: CD68+PDL1-IDO1- Area (μm²)"='area_pdl1neg_ido1neg_cd68pos_tumour',
  "Tumour: PDL1-IDO1-CD68- Area (μm²)"='area_pdl1neg_ido1neg_cd68neg_tumour',
  "Stroma: PDL1+IDO1+CD68- Area (μm²)"='area_pdl1pos_ido1pos_cd68neg_stroma',
  "Stroma: PDL1+IDO1-CD68- Area (μm²)"='area_pdl1pos_ido1neg_cd68neg_stroma',
  "Stroma: PDL1-IDO1+CD68- Area (μm²)"='area_pdl1neg_ido1pos_cd68neg_stroma',
  "Stroma: PDL1+IDO1-CD68+ Area (μm²)"='area_pdl1pos_ido1neg_cd68pos_stroma',
  "Stroma: PDL1-IDO1+CD68+ Area (μm²)"='area_pdl1neg_ido1pos_cd68pos_stroma',
  "Stroma: PDL1+IDO1+CD68+ Area (μm²)"='area_pdl1pos_ido1pos_cd68pos_stroma',
  "Stroma: CD68+PDL1-IDO1- Area (μm²)"='area_pdl1neg_ido1neg_cd68pos_stroma',
  "Stroma: PDL1-IDO1-CD68- Area (μm²)"='area_pdl1neg_ido1neg_cd68neg_stroma'
)

ar_area_22004_labeled <- ar_area_22004_labeled %>% 
  rename_with(~ colname_map_area2[.], .cols = names(colname_map_area2)) %>%
  filter(ifelse(is.na(notes), TRUE, !str_detect(notes, pattern = "(Exclude|exclude|discard)"))) %>%
  mutate(acc_num=str_replace(acc_num, " \\(.*", "")) %>%
  select(c(tma, sector, row, col, acc_num, block, unname(colname_map_area2)))
```

### Merged

```{r}
ar_pheno_all <- ar_pheno_100056_labeled %>%
  bind_rows(ar_pheno_22004_labeled)

ar_area_all <- ar_area_100056_labeled %>%
  bind_rows(ar_area_22004_labeled)
```

```{r}
ar_output_path <- here::here("data", "processed", "ar_pheno_labeled.tsv")
ar_area_output_path <- here::here("data", "processed", "ar_area_labeled.tsv")

write_tsv(ar_pheno_all, file = ar_output_path)
write_tsv(ar_area_all, file = ar_area_output_path)
```




# Angiogenesis

We'll import angiogenesis results for the purposes of looking at vessel density, as our ability to assess the other markers is limited due to significant batch effects. 

```{r}
angio_dir <- here::here("data", "raw", "Angio Panel")

angio_map_files <- file.path(angio_dir, "Angio SectorMaps", 
                             c("10-005 G3 Endometrioid Endometrial Carcinoma.xls",
                               "10-006 High-Grade Serous Endometrial Carcinoma.xls",
                               "22-004 P53 Abnormal Endometrial Carcinoma Tumours (mixed sites).xlsx")
)

sectormaps <- list(
  '10-005'=read_excel(angio_map_files[1], sheet = "SectorMap-HALO", col_names = FALSE, skip = 2),
  '10-006'=read_excel(angio_map_files[2], sheet = "SectorMap-HALO2", col_names = FALSE, skip = 2),
  '22-004A'=read_excel(angio_map_files[3], sheet = "SectorMap A-HALO", col_names = FALSE, skip = 2),
  '22-004B'=read_excel(angio_map_files[3], sheet = "SectorMap B-HALO", col_names = FALSE, skip = 2)
)

core_ids <- list(
  '10-005'=read_excel(angio_map_files[1], sheet = "CoreID", skip = 2),
  '10-006'=read_excel(angio_map_files[2], sheet = "CoreID", skip = 2),
  '22-004A'=read_excel(angio_map_files[3], sheet = "Core ID", skip = 2),
  '22-004B'=read_excel(angio_map_files[3], sheet = "Core ID", skip = 2)
)
```

```{r}
sectormaps_long <- lapply(names(sectormaps), function(x) {
  convert_sectormap_to_long(sectormaps[[x]]) %>%
    mutate(tma=x) %>%
    select(tma, everything())
}) %>% bind_rows()

core_ids_long <- lapply(names(core_ids), function(x) {
  df <- core_ids[[x]] %>%
    mutate(tma=x) %>%
    rename(core=CoreID) %>%
    select(tma, everything())
  colnames(df) <- str_replace(colnames(df), "( )*#$", "")
  return(df)
}) %>% bind_rows()

# Sample map, in long format, derived from the weird TMA maps
sample_map <- sectormaps_long %>% 
  inner_join(core_ids_long) %>%
  rename(acc_num=Accession, block=Block, backup=Backup) %>%
  select(tma, sector, core, row, col, acc_num, block, backup) %>%
  dplyr::mutate(
    row = as.character(row),
    col = as.character(col)
  )
```


```{r}
angio_data_dir <- file.path(angio_dir, "Final FL 10-005 10-006 22-004A-B Angio")
angio_data_files <- dir_ls(angio_data_dir, regexp = ".xls(x)?")

angio_data_files <- angio_data_files[!str_detect(angio_data_files, "ERG1")]

angio_data <- lapply(angio_data_files, function(fp) {
  sheetnames <- excel_sheets(fp)
  
  region_dfs <- lapply(sheetnames, function(x) {
    df <- read_excel(fp, sheet = x, na = c("", "x", "tk"))
    df <- df %>%
      rename(row=`TMA Row`,
             col=`TMA Column`,
             tma=`TMA Block Name`,
             notes=Notes) %>%
      select(-c(`Image Stain`, `Area Analyzed (μm²)`)) %>%
      mutate(tma=str_extract(x, "[0-9]{2}\\-[0-9]{3}[A-Z]?"),
             row=as.character(row),
             col=as.character(col))
    
    return(df)
  }) 
  
  region_dfs_merged <- region_dfs %>% reduce(inner_join)
}) %>% bind_rows()

angio_data_labeled <- angio_data %>%
  inner_join(sample_map) %>%
  select(tma, acc_num, row, col, notes, everything()) %>%
  arrange(tma, row, col, acc_num)
```

```{r}
colname_map <- c(
  "Tumour Area (mm²)"='area_tumour_mm',
  "Stroma Area (mm²)"='area_stroma_mm',
  "Vessel Area (mm²)"='area_vessel_mm'
)

angio_data_labeled <- angio_data_labeled %>% 
  rename_with(~ colname_map[.], .cols = names(colname_map)) %>%
  # This rule shouldn't apply
  filter(ifelse(is.na(notes), TRUE, !str_detect(notes, pattern = "(Exclude|exclude|discard)"))) %>%
  # This rule shouldn't apply
  mutate(acc_num=str_replace(acc_num, " \\(.*", "")) %>%
  select(c(tma, sector, row, col, acc_num, block, unname(colname_map)))
```

```{r}
angio_output_path <- here::here("data", "processed", "angio_labeled.tsv")

write_tsv(angio_data_labeled, file = angio_output_path)
```




