---
title: "AR analysis"
output: 
  html_document:
      code_folding: hide
      toc: true
      toc_float: true
---


```{r global_chunk_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, warning = FALSE, message = FALSE, cache = TRUE, cache.lazy = FALSE, fig.width = 8, fig.height = 4.5)

here::i_am("notebooks/ar_analysis.Rmd")
```

```{r, echo = TRUE}
library(tidyverse)
library(readxl)
library(ComplexHeatmap)
library(textclean)
library(survival)
library(survminer)
library(circlize)
library(glmnet)
library(reticulate)
library(ggupset)
library(ggforce)
library(RColorBrewer)
library(car)
library(here)
```



# Data

```{r}
ar_area_file <- here("data", "AR_area_labeled.tsv")
ar_pheno_file <- here("data", "AR_pheno_labeled.tsv")
clinical_data_file <- here("data", "clinical_data.tsv")

ar_area_raw <- read_tsv(ar_area_file)
ar_pheno_raw <- read_tsv(ar_pheno_file)

clinical_data <- read_tsv(clinical_data_file)
```


```{r}
colname_map <- c(
  '% Tumour'='area_tumour_pct',
  '% Stroma'='area_stroma_pct',
  "Classified Area (mm²)"='area_tissue_mm',
  "Tumour Area (mm²)"='area_tumour_mm',
  "Stroma Area (mm²)"='area_stroma_mm',
  "Tumour: CD68+PDL1-IDO1- Cells"='count_mac_pdl1neg_ido1neg_tumour',
  "Stroma:CD68+PDL1-IDO1- Cells"='count_mac_pdl1neg_ido1neg_stroma',
  "Tumour: CD68+PDL1+IDO1- Cells"='count_mac_pdl1pos_ido1neg_tumour',
  "Stroma:  CD68+PDL1+IDO1- Cells"='count_mac_pdl1pos_ido1neg_stroma',
  "Tumour: CD68+PDL1+IDO1+ Cells"='count_mac_pdl1pos_ido1pos_tumour',
  "Stroma: CD68+PDL1+IDO1+ Cells"='count_mac_pdl1pos_ido1pos_stroma',
  "Tumour: CD68+PDL1-IDO1+ Cells"='count_mac_pdl1neg_ido1pos_tumour',
  "Stroma: CD68+PDL1-IDO1+ Cells"='count_mac_pdl1neg_ido1pos_stroma',
  "Tumour: CD8+PD1- Cells"='count_tcyto_pd1neg_tumour',
  "Stroma: CD8+PD1- Cells"='count_tcyto_pd1neg_stroma',
  "Tumour: CD8+PD1+ Cells"='count_tcyto_pd1pos_tumour',
  "Stroma: CD8+PD1+ Cells"='count_tcyto_pd1pos_stroma',
  "Tumour: CD8-PD1+ Cells"='count_tcd8neg_pd1pos_tumour',
  "Stroma: CD8-PD1+ Cells"='count_tcd8neg_pd1pos_stroma'
)

ar_pheno_data <- ar_pheno_raw %>% 
  rename_with(~ colname_map[.], .cols = names(colname_map)) %>%
  # This rule shouldn't apply
  filter(ifelse(is.na(notes), TRUE, !str_detect(notes, pattern = "(Exclude|exclude)"))) %>%
  # This rule shouldn't apply
  mutate(acc_num=str_replace(acc_num, " \\(.*", "")) %>%
  select(c("tma", "acc_num", unname(colname_map)))

ar_pheno_long <- ar_pheno_data %>%
  pivot_longer(cols = all_of(colnames(ar_pheno_data)[str_detect(colnames(ar_pheno_data), "^(area|count)")]), 
               names_to = "variable", 
               values_to = "value")
```

```{r}
area_distributions <- ar_pheno_long %>%
  filter(str_detect(variable, "^area")) %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 10, fill = "lightblue", color = "black") +
  facet_wrap(vars(variable), scales = "free") +
  labs(title = "Variables") + 
  theme_pubclean() + 
  ggtitle("Region area per core")

count_distributions <- ar_pheno_long %>%
  filter(str_detect(variable, "^count")) %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  scale_x_continuous(trans = 'log1p', breaks = c(0, 1, 10, 100, 1000, 10000),
                     labels = c("0", "1", "10", "100", "1000", "10000")) + 
  facet_wrap(vars(variable), scales = "free") +
  labs(title = "Variables") + 
  theme_pubr() + 
  ggtitle("Count per core")
```

## Region area distribution

```{r, fig.width=6, fig.height=4}
area_distributions
```

## Count distribution

```{r, fig.width=10, fig.height=6}
count_distributions
```

```{r}
ar_counts_long <- ar_pheno_data %>%
  pivot_longer(cols = colnames(ar_pheno_data)[startsWith(colnames(ar_pheno_data), "count")],
               names_to = "variable", 
               values_to = "value") %>%
  mutate(area_region_mm=ifelse(str_detect(variable, "tumour"),
                               area_tumour_mm,
                               area_stroma_mm))

ar_counts_sum_long <- ar_counts_long %>%
  group_by(acc_num, variable) %>%
  summarise(
    var=var(value/area_region_mm),
    var_count=var(value),
    count=round(sum(value), 0),
    value=sum(value)/sum(area_region_mm),
    area=sum(area_region_mm)
  ) %>%
  ungroup() %>%
  mutate(variable = str_replace(variable, "^count", "dens"),
         unique_id=acc_num) %>%
  select(unique_id, everything())
```

## Mean-variance

```{r, fig.width=6, fig.height=4}
ggplot(ar_counts_sum_long, aes(x=count, y=var_count)) + 
  geom_point(alpha = 0.3) +
  stat_smooth(method="loess", colour="blue", alpha = 0.3) + 
  facet_wrap(~ variable) + 
  theme_minimal() + 
  scale_x_continuous(trans = "log1p", breaks = c(0, 1, 10, 100, 1000, 10000)) + 
  scale_y_continuous(trans = "log1p", breaks = c(0, 1, 10, 100, 1000, 10000)) + 
  geom_abline(slope=1, colour='red', alpha = 0.3) + 
  xlab("Count") + 
  ylab("Variance")
```

```{r}
ar_densities <- ar_counts_sum_long %>%
  mutate(value = log1p(value)) 

ar_densities_wide <- ar_densities %>%
  select(-c(count, area, var, var_count)) %>%
  pivot_wider(names_from = variable, values_from = value)

ar_densities_heatmap_input <- ar_densities_wide %>% select(colnames(ar_densities_wide)[str_detect(colnames(ar_densities_wide), "^dens")]) %>% as.matrix
rownames(ar_densities_heatmap_input) <- ar_densities_wide$unique_id
```

```{r}
group_by_cols <- setdiff(colnames(clinical_data), c("tma", "core_id"))

clinical_data_unique <- clinical_data %>% 
  group_by_at(group_by_cols) %>%
  summarise(across(everything(), ~ paste(., collapse = ","))) %>%
  ungroup()

clinical_anno <- clinical_data_unique %>%
  select(acc_num, tma, grade_rev, hist_rev, neoadj, eclass2_ngs) %>%
  mutate(unique_id=acc_num) %>%
  column_to_rownames("unique_id") %>%
  select(-c(acc_num))
```


```{r, fig.width = 12, fig.height = 5}
input_mat <- scale(ar_densities_heatmap_input %>% na.omit)

## REMOVE THIS IF YOU WANT THE ENTIRE COHORT 
#input_mat <- input_mat[which(row_anno$eclass2_ngs == "p53abn"),] %>% scale

row_hclust <- hclust(dist(input_mat), method = "ward.D2")
col_hclust <- hclust(dist(t(input_mat)), method = "ward.D2")

ar_clusters <- cutree(row_hclust, 3)
names(ar_clusters) <- rownames(input_mat)

outcome_clusters <- read_tsv(here::here("results", "outcome_clusters_long.tsv"))
tilclust_assignments <- outcome_clusters %>% 
  filter(outcome == "pfs") %>%
  dplyr::select(acc_num, clust) %>%
  dplyr::mutate(clust = factor(clust, levels = c("0", "1")))
  unique

tilclust_assignments_rn <- tilclust_assignments %>% 
  column_to_rownames("acc_num")

pca_res <- prcomp(input_mat)

row_anno <- clinical_anno[rownames(input_mat),]
row_anno$ar_clusters <- factor(unname(ar_clusters))
hist_mapping <- c("endometrioid (squamous)" = "endometrioid", 
                  "mixed endometrioid and serous" = "mixed", 
                  "mixed serous and moderately differentiated" = "mixed",
                  "undifferentiated" = "undiff/dediff",
                  "dedifferentiated" = "undiff/dediff")
row_anno <- row_anno %>%
  mutate(hist_rev = case_when(
    hist_rev %in% names(hist_mapping) ~ hist_mapping[hist_rev],
    TRUE ~ hist_rev 
  ))
row_anno <- row_anno %>%
  cbind(pca_res$x[,1:4])
row_anno$tilclust <- tilclust_assignments_rn[rownames(row_anno),]

ha <- rowAnnotation(df = row_anno, annotation_width = 0.02, col = list(
  PC1=colorRamp2(c(min(row_anno$PC1), 0, max(row_anno$PC1)), c("blue", "white", "red")),
  PC2=colorRamp2(c(min(row_anno$PC2), 0, max(row_anno$PC2)), c("blue", "white", "red")),
  PC3=colorRamp2(c(min(row_anno$PC3), 0, max(row_anno$PC3)), c("blue", "white", "red")),
  PC4=colorRamp2(c(min(row_anno$PC4), 0, max(row_anno$PC4)), c("blue", "white", "red"))))

density_heatmap <- Heatmap(input_mat,
                           na_col = "gray",
                           column_names_gp = gpar(fontsize = 6), 
                           cluster_rows = row_hclust,
                           cluster_columns = col_hclust,
                           column_names_rot = 60, 
                           show_row_names = FALSE) + ha


draw(density_heatmap, heatmap_legend_side = "right")
```
```{r, fig.width = 12, fig.height = 5}
density_heatmap <- Heatmap(input_mat,
                           na_col = "gray",
                           column_names_gp = gpar(fontsize = 6), 
                           cluster_rows = TRUE,
                           cluster_columns = TRUE,
                           column_names_rot = 60, 
                           show_row_names = FALSE,
                           row_split = row_anno$tilclust) + ha


draw(density_heatmap, heatmap_legend_side = "right")
```

```{r, fig.width = 10, fig.height = 10}
ar_densities_clusts <- ar_densities %>%
  left_join(tilclust_assignments)

ggplot(ar_densities_clusts, aes(x=clust, y=value)) + 
  geom_boxplot(width = 0.5, outlier.size = -1) + 
  geom_point(position = position_jitter(width = 0.2, height = 0), alpha = 0.3) + 
  theme_pubr() + 
  stat_compare_means(vjust = 0.7, size = 3) + 
  facet_wrap(~ variable) 
```



# OLD AREA ANALYSIS -- NOT USEFUL 

```{r}
ar_area <- ar_area_raw %>%
  filter(is.na(notes) | str_detect(notes, "((?i)discard|(?i)exclude)", negate = TRUE)) %>%
  mutate(acc_num=str_replace(acc_num, "-(0+)", "-")) %>%
  mutate(acc_num=str_replace(acc_num, " \\(.*", ""))

colnames(ar_area) <- str_replace(colnames(ar_area), " \\(Opal [0-9]+\\)", "")
```

```{r}
ar_area %>% dim
```

```{r}
ar_area %>% colnames
```

```{r}
area_features <- c(
  "PanCK+ Area (μm²)",
  "PDL1 Area (μm²)", "PDL1 Weak Area (μm²)", "PDL1 Moderate Area (μm²)", "PDL1 Strong Area (μm²)",
  "IDO1 Area (μm²)", "IDO1 Weak Area (μm²)", "IDO1 Moderate Area (μm²)", "IDO1 Strong Area (μm²)",
  "CD68 Area (μm²)", "CD68 Weak Area (μm²)", "CD68 Moderate Area (μm²)", "CD68 Strong Area (μm²)",
  "CD8 Area (μm²)", "CD8 Weak Area (μm²)", "CD8 Moderate Area (μm²)", "CD8 Strong Area (μm²)",
  "PD1 Area (μm²)", "PD1 Weak Area (μm²)", "PD1 Moderate Area (μm²)", "PD1 Strong Area (μm²)",
  "PDL1+IDO1+CD68- Area (μm²)", "PDL1+IDO1-CD68- Area (μm²)", "PDL1-IDO1+CD68- Area (μm²)", 
  "PDL1+IDO1-CD68+ Area (μm²)", "PDL1-IDO1+CD68+ Area (μm²)", "PDL1+IDO1+CD68+ Area (μm²)",
  "CD8+PD1+ Area (μm²)", "CD8-PD1+ Area (μm²)", "CD8+PD1- Area (μm²)", "CD68+IDO1-PDL1- Area (μm²)", "PDL1-IDO1-CD68- Area (μm²)"
)

intensity_features <- colnames(ar_area)[str_detect(colnames(ar_area), "^(Tumour|Stroma)") & str_detect(colnames(ar_area), "Intensity")]
```

```{r}
ar_area_selected <- ar_area %>%
  select(c(
    "tma", "acc_num", "sector", "row", "col", "Algorithm Name", "Classified Area (mm²)", "Glass Area (mm²)", "Other Area (mm²)",
    "Stroma Area (mm²)", "Tumour Area (mm²)",
    apply(expand.grid(c("Tumour", "Stroma"), area_features), 1, paste, collapse=": "),
    intensity_features
  )) %>%
  rename(algorithm=`Algorithm Name`)
```

```{r}
## Summarize the areas first, as those only require summation
ar_areas_long <- ar_area_selected %>%
  gather(variable, value, -c("tma", "acc_num", "sector", "row", "col", "algorithm")) %>%
  filter(str_detect(variable, "Area")) %>%
  mutate(value = replace_na(value, 0)) %>%
  ## Homogenize area units
  mutate(value = ifelse(str_detect(variable, "μm²"), value/10^6, value)) %>%
  mutate(variable = str_replace_all(variable, "μm²", "mm²"))

ar_areas_summarized_long <- ar_areas_long %>%
  group_by(tma, acc_num, algorithm, variable) %>%
  summarise(value=sum(value)) %>%
  ungroup()


## Get intensity variables
ar_intensities <- ar_area_selected %>% 
  gather(variable, value, -c("tma", "acc_num", "sector", "row", "col", "algorithm")) %>%
  filter(str_detect(variable, "Intensity")) %>%
  mutate(value = replace_na(value, 0)) %>%
  rename(intensity=value)

## Parse corresponding area string
ar_intensities <- ar_intensities %>%
  mutate(corresponding_area=str_replace(trimws(str_replace(variable, "[a-zA-Z0-9\\+\\-]+ Average (Positive )?Intensity", "")), "\\:$", "")) %>%
  mutate(corresponding_area = paste0(corresponding_area, " Area (mm²)"))

ar_intensities_areas <- ar_intensities %>%
  left_join(ar_areas_long %>%
              rename(corresponding_area=variable, area=value))

## Summarize over cores for the same tma and acc_num
ar_average_intensities_by_tma_sample <- ar_intensities_areas %>%
  group_by(tma, acc_num, variable, corresponding_area) %>%
  summarise(intensity=sum(intensity*area)/sum(area),
            area=sum(area)) %>%
  ungroup()

ar_average_intensities_by_sample <- ar_intensities_areas %>%
  group_by(acc_num, variable, corresponding_area) %>%
  summarise(intensity=sum(intensity*area)/sum(area),
            area=sum(area)) %>%
  ungroup()
```

## TMA/Sample heatmap

```{r}
ar_average_intensities_by_tma_sample_wide <- ar_average_intensities_by_tma_sample %>%
  select(-c(corresponding_area, area)) %>%
  spread(key = variable, value = intensity) %>%
  mutate(unique_id=paste0(tma, "_", acc_num)) %>%
  select(-c(tma, acc_num)) %>%
  select(unique_id, everything())

ar_average_intensities_by_tma_sample_matrix <- ar_average_intensities_by_tma_sample_wide %>%
  column_to_rownames("unique_id") %>% 
  scale

ar_average_intensities_by_tma_sample_matrix_filtered <- ar_average_intensities_by_tma_sample_matrix[rowSums(is.na(ar_average_intensities_by_tma_sample_matrix)) != ncol(ar_average_intensities_by_tma_sample_matrix),]

ar_average_intensities_by_tma_sample_matrix_filtered <- ar_average_intensities_by_tma_sample_matrix_filtered[,colSums(is.na(ar_average_intensities_by_tma_sample_matrix_filtered)) != nrow(ar_average_intensities_by_tma_sample_matrix_filtered)]
  
row_annos <- data.frame(
  unique_id=rownames(ar_average_intensities_by_tma_sample_matrix_filtered)
) %>%
  mutate(tma=str_extract(unique_id, "^[0-9]{2}\\-[0-9]{3}([A-Z])?")) %>%
  column_to_rownames("unique_id")

col_annos <- data.frame(
  name=colnames(ar_average_intensities_by_tma_sample_matrix_filtered)
) %>%
  mutate(region=str_extract(name, "^[A-Za-z]+")) %>%
  column_to_rownames("name")

ra <- rowAnnotation(df = row_annos)
ca <- columnAnnotation(df = col_annos)


ar_intensity_heatmap <- Heatmap(scale(ar_average_intensities_by_tma_sample_matrix_filtered),
                                na_col = "gray",
                                column_names_gp = gpar(fontsize = 6), 
                                cluster_rows = TRUE,
                                cluster_columns = TRUE,
                                column_names_rot = 80, 
                                show_row_names = FALSE) + ra


draw(ar_intensity_heatmap, heatmap_legend_side = "right")
```

So there are significant batch effects here, but hopefully the thresholded data is good. 

## Areas

```{r}
general_area_variables <- c("Classified Area (mm²)", "Glass Area (mm²)", "Other Area (mm²)",
                         "Stroma Area (mm²)", "Tumour Area (mm²)")

ar_general_areas <- ar_areas_long %>%
  filter(variable %in% general_area_variables) %>%
  mutate(parent_area=str_extract(variable, "^(Tumour|Stroma)"))

ar_specific_areas <- ar_areas_long %>% 
  filter(!variable %in% general_area_variables) %>%
  mutate(parent_area=str_extract(variable, "^(Tumour|Stroma)")) %>%
  left_join(
    ar_general_areas %>%
      select(tma, acc_num, sector, row, col, parent_area, value) %>%
      filter(!is.na(parent_area)) %>%
      rename(parent_value=value)
  ) %>%
  rename(area=variable)

# frac_value: Fractional area within parent region
ar_specific_areas <- ar_specific_areas %>%
  mutate(frac_value=value/parent_value,
         area=str_replace(area, " Area \\(mm²\\)", ""))
```


```{r, fig.width = 10, fig.height = 8}
ggplot(ar_specific_areas, aes(x=frac_value, fill=tma)) + 
  geom_density(alpha = 0.3) + 
  facet_wrap(~ area, scales = "free") + 
  theme_pubr() + 
  scale_x_log10() + 
  xlab("Fraction of parent area")
```

Those look reasonable. 

## Functional areas (i.e. non-phenotyped areas)

```{r}
# Get the ACTUAL biological columns of interest
ar_specific_areas_functional <- ar_areas_long %>%
  mutate(area=trimws(str_replace(variable, " Area \\(mm²\\)", ""))) %>%
  select(-c(variable)) %>%
  group_by(acc_num) %>%
  summarise(
    # CD68+ cells
    `Tumour: CD68+ PDL1+IDO1+`=sum(value[area == "Tumour: PDL1+IDO1+CD68+"])/sum(value[area == "Tumour: CD68"]),
    `Tumour: CD68+ PDL1-IDO1+`=sum(value[area == "Tumour: PDL1-IDO1+CD68+"])/sum(value[area == "Tumour: CD68"]),
    `Tumour: CD68+ PDL1+IDO1-`=sum(value[area == "Tumour: PDL1+IDO1-CD68+"])/sum(value[area == "Tumour: CD68"]),
    `Tumour: CD68+ PDL1-IDO1-`=sum(value[area == "Tumour: CD68+IDO1-PDL1-"])/sum(value[area == "Tumour: CD68"]),
    `Stroma: CD68+ PDL1+IDO1+`=sum(value[area == "Stroma: PDL1+IDO1+CD68+"])/sum(value[area == "Stroma: CD68"]),
    `Stroma: CD68+ PDL1-IDO1+`=sum(value[area == "Stroma: PDL1-IDO1+CD68+"])/sum(value[area == "Stroma: CD68"]),
    `Stroma: CD68+ PDL1+IDO1-`=sum(value[area == "Stroma: PDL1+IDO1-CD68+"])/sum(value[area == "Stroma: CD68"]),
    `Stroma: CD68+ PDL1-IDO1-`=sum(value[area == "Stroma: CD68+IDO1-PDL1-"])/sum(value[area == "Stroma: CD68"]),
    # CD8+ cells not possible to do
    # Tumour cells
    `Tumour: CD68- PDL1+IDO1+`=sum(value[area == "Tumour: PDL1+IDO1+CD68-"])/sum(value[area == "Tumour"] - value[area == "Tumour: CD68"]),
    `Tumour: CD68- PDL1-IDO1+`=sum(value[area == "Tumour: PDL1-IDO1+CD68-"])/sum(value[area == "Tumour"] - value[area == "Tumour: CD68"]),
    `Tumour: CD68- PDL1+IDO1-`=sum(value[area == "Tumour: PDL1+IDO1-CD68-"])/sum(value[area == "Tumour"] - value[area == "Tumour: CD68"]),
    `Tumour: CD68- PDL1-IDO1-`=sum(value[area == "Tumour: PDL1-IDO1-CD68-"])/sum(value[area == "Tumour"] - value[area == "Tumour: CD68"]),
    # Stromal cells
    `Stroma: CD68- PDL1+IDO1+`=sum(value[area == "Stroma: PDL1+IDO1+CD68-"])/sum(value[area == "Stroma"] - value[area == "Stroma: CD68"]),
    `Stroma: CD68- PDL1-IDO1+`=sum(value[area == "Stroma: PDL1-IDO1+CD68-"])/sum(value[area == "Stroma"] - value[area == "Stroma: CD68"]),
    `Stroma: CD68- PDL1+IDO1-`=sum(value[area == "Stroma: PDL1+IDO1-CD68-"])/sum(value[area == "Stroma"] - value[area == "Stroma: CD68"]),
    `Stroma: CD68- PDL1-IDO1-`=sum(value[area == "Stroma: PDL1-IDO1-CD68-"])/sum(value[area == "Stroma"] - value[area == "Stroma: CD68"]),
  ) %>%
  ungroup()

# ar_specific_areas_functional <- ar_specific_areas_functional %>%
#   mutate(unique_id = paste0(tma, "_", acc_num)) %>%
#   select(-c(tma, acc_num)) %>%
#   select(unique_id, everything())
```

```{r}
# Read in TIL clusters and TIL densities
til_densities_output_file <- "../results/til_densities_output.tsv"

# NB clusters
outcome_clusters_long_file <- "../results/outcome_clusters_long.tsv"

til_densities_output <- read_tsv(til_densities_output_file)
outcome_clusters_long <- read_tsv(outcome_clusters_long_file)
```

```{r}
ar_specific_areas_functional_matrix <- ar_specific_areas_functional %>%
  column_to_rownames("acc_num")
  
row_annos <- outcome_clusters_long %>%
  filter(outcome == "os") %>%
  select(acc_num, clust) %>%
  mutate(clust = factor(clust)) %>%
  column_to_rownames("acc_num")

row_annos <- row_annos[rownames(ar_specific_areas_functional_matrix),,drop=FALSE]
rownames(row_annos) <- rownames(ar_specific_areas_functional_matrix)

col_annos <- data.frame(
  name=colnames(ar_specific_areas_functional_matrix)
) %>%
  mutate(region=str_extract(name, "^[A-Za-z]+"),
         parent_celltype=str_extract(name, "CD68[\\+\\-]")) %>%
  column_to_rownames("name")

clust_color_map <- c('0'='#60B2E5',
                     '1'='#A4031F')

ra <- rowAnnotation(df = row_annos,
                    col = list(clust = clust_color_map))
ca <- columnAnnotation(df = col_annos)


ar_functional_heatmap <- Heatmap(ar_specific_areas_functional_matrix %>% scale,
                                na_col = "gray",
                                column_names_gp = gpar(fontsize = 6), 
                                cluster_rows = TRUE,
                                cluster_columns = TRUE,
                                column_names_rot = 80, 
                                show_row_names = FALSE,
                                top_annotation = ca, 
                                row_split = row_annos$clust) + ra


draw(ar_functional_heatmap, heatmap_legend_side = "right")
```

I plotted the above with TMA as a stratifying variable too, and the batch effects appear to be insignificant. 
